---
title: "Data Story 2"
author: "Walker Briggs"
date: "2026-02-26"
output: ioslides_presentation
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE }
library(dplyr)
library(readr)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE }

# Loading in my data from Our World in Data

url <- 'https://ourworldindata.org/explorers/poverty-explorer.csv?v=1&csvType=full&useColumnShortNames=true&Indicator=Share+in+poverty&Poverty+line=%243+per+day%3A+International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+consumption+surveys&Show+breaks+between+less+comparable+surveys=false'
poverty <- read_csv(url)
# poverty %>% head

```

```{r, echo=FALSE, warning=FALSE, message=FALSE }

# Loading in my data from Our World in Data

url <- 'https://ourworldindata.org/grapher/prevalence-of-undernourishment.csv?v=1&csvType=full&useColumnShortNames=true'
food <- read_csv(url)
# food %>% head

```

```{r, echo=FALSE, warning=FALSE, message=FALSE }

# Loading in my data from Our World in Data

url <- 'https://ourworldindata.org/grapher/proportion-using-safely-managed-drinking-water.csv?v=1&csvType=full&useColumnShortNames=true'
water <- read_csv(url)
# water %>% head

```

```{r, echo=FALSE, warning=FALSE, message=FALSE }

# Loading in my data from Our World in Data

url <- 'https://ourworldindata.org/grapher/deaths-in-armed-conflicts-by-country.csv?v=1&csvType=full&useColumnShortNames=true'
war <- read_csv(url)
# war %>% head

```

```{r, echo=FALSE, warning=FALSE, message=FALSE }
library(gapminder)
library(dplyr)
data(gapminder)
#gapminder %>% head
```


## Main Focus: Extreme Poverty

According to Our World In Data, extreme poverty is defined as living below
the international poverty line of $3 per day.


### **Key Objectives**:

- *Objective 1*: Exploring global extreme poverty rates
- *Objective 2*: Using malnourishment data
- *Objective 3*: Incorporating accessibility to clean drinking water data
- *Objective 4*: Is there a connection between extreme poverty and armed conflict?



## Sustainable Development Goals
- SDG 1: **No Poverty**

&nbsp;

- SDG 2: **Zero Hunger**

&nbsp;

- SDG 6: **Clean Water and Sanitation**

&nbsp;
&nbsp;

*All data used for this data story was found from Our World In Data.*

## Exploring The Data

```{r, echo=FALSE, warning=FALSE, message=FALSE }

library(dplyr)
library(readr)
library(ggplot2)

poverty <- read_csv("https://ourworldindata.org/explorers/poverty-explorer.csv?v=1&csvType=full&useColumnShortNames=true&Indicator=Share+in+poverty&Poverty+line=%243+per+day%3A+International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+consumption+surveys&Show+breaks+between+less+comparable+surveys=false")

poverty %>%
  filter(country %in% c("Australia", "Russia", "Canada",
                       "United States", "United Kingdom")) %>%
  ggplot(aes(x = year,
             y = `share_below_$3_a_day`,
             color = country)) +
   geom_path(size=1) +
  labs(
    title = "Countries Earning More Per Capita",
    x = "Year",
    y = "Earning below $3 per day (% of population)",
    color = "Country")

```

## Exploring The Data

```{r, echo=FALSE, warning=FALSE, message=FALSE }

library(dplyr)
library(readr)
library(ggplot2)

poverty <- read_csv("https://ourworldindata.org/explorers/poverty-explorer.csv?v=1&csvType=full&useColumnShortNames=true&Indicator=Share+in+poverty&Poverty+line=%243+per+day%3A+International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+consumption+surveys&Show+breaks+between+less+comparable+surveys=false")

poverty %>%
  filter(country %in% c("Democratic Republic of Congo", "Mozambique", "Malawi",
                       "Burundi", "Zambia")) %>%
  ggplot(aes(x = year,
             y = `share_below_$3_a_day`,
             color = country)) +
   geom_path(size=1) +
  labs(
    title = "Countries Living In Extreme Poverty",
    x = "Year",
    y = "Earning below $3 per day (% of population)",
    color = "Country")

```

## Exploring The Data

```{r, echo=FALSE, warning=FALSE, message=FALSE }

library(dplyr)
library(readr)
library(ggplot2)

food <- read_csv("https://ourworldindata.org/grapher/prevalence-of-undernourishment.csv?v=1&csvType=full&useColumnShortNames=true")

food %>%
  filter(entity %in% c("Democratic Republic of Congo", "Mozambique", "Malawi",
                       "Burundi", "Zambia")) %>%
  ggplot(aes(x = year,
             y = `_2_1_1_prevalence_of_undernourishment__000000000024000__value__006121__percent`,
             color = entity)) +
   geom_area(size=1) +
  labs(
    title = "Share of People Who Are Undernourished",
    x = "Year",
    y = "Percent of Population Undernourished",
    color = "Country")

```

## Exploring The Data

```{r, echo=FALSE, warning=FALSE, message=FALSE }

library(dplyr)
library(readr)
library(ggplot2)

water <- read_csv("https://ourworldindata.org/grapher/proportion-using-safely-managed-drinking-water.csv?v=1&csvType=full&useColumnShortNames=true")

water %>%
  filter(entity %in% c("Democratic Republic of Congo", "Mozambique", "Malawi",
                       "Burundi", "Zambia")) %>%
  ggplot(aes(x = year,
             y = `wat_sm__residence_total`,
             color = entity)) +
   geom_area(size=1) +
  labs(
    title = "Share of People Using Managed/Safe Drinking Water",
    x = "Year",
    y = "Percent of Population Able to Access Clean Drinking Water",
    color = "Country")

```

## Exploring The Data

```{r, echo=FALSE, warning=FALSE, message=FALSE }

library(dplyr)
library(readr)
library(ggplot2)

war <- read_csv("https://ourworldindata.org/grapher/deaths-in-armed-conflicts-by-country.csv?v=1&csvType=full&useColumnShortNames=true")

war %>%
  filter(entity %in% c("Democratic Republic of Congo", "Mozambique", "Malawi",
                       "Burundi", "Zambia")) %>%
  ggplot(aes(x = year,
             y = `number_deaths_ongoing_conflicts__conflict_type_all`,
             color = entity)) +
   geom_path(size=1) +
  labs(
    title = "Deaths as a Result of Armed Conflict",
    x = "Year",
    y = "Number of Deaths",
    color = "Country")

```

## Conclusion
- Clear connection between extreme poverty and malnourishment.
- People using managed/safe drinking water has been increasing since the 
early 2000s for most impoverished countries.
- The connection between extreme poverty and deaths as a result of armed
conflict is more unclear. These regions experiencing extreme poverty have
had armed conflict situations yet it may not have as direct of an impact
on poverty. This is a misconception that many people, including myself
fell victim to.